---
title: "Assignment 1"
author: "Kathirmani Sukumar"
date: "September 26, 2017"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(knitr)
parliament = read.csv('d:/datasets/parliament.csv')
names(parliament)
```
### Question 1
Identify year wise total number of electors and voters? Also compute percentage of voters?
```{r warning=FALSE, message=FALSE}
summary = parliament %>% group_by(YEAR, STATE, PC) %>% summarise(VOTES=sum(VOTES, na.rm=TRUE), ELECTORS=first(ELECTORS))
yearwise = summary %>% group_by(YEAR) %>% summarise(ELECTORS=sum(ELECTORS, na.rm=TRUE), VOTES=sum(VOTES, na.rm=TRUE)) %>% mutate(PERC=VOTES/ELECTORS*100)
kable(yearwise)
```

### Question 2
Identify year wise percentage of male and female candidates?
```{r warning=FALSE, message=FALSE}
summary = as.data.frame.matrix(table(parliament$YEAR, parliament$SEX))
summary$Total = rowSums(summary)
summary$male_per = summary$M / summary$Total * 100
summary$female_per = summary$F / summary$Total * 100
kable(summary)
```

### Question 3
Compute party wise total number of candidates won so far? Display top ten parties alone
```{r}
winning_rows = filter(parliament, Position==1)
summary_3 = winning_rows %>% group_by(PARTY) %>% summarise(Total=n()) %>% arrange(-Total)
head(summary_3, 10)
```

### Question 4
Identify year wise which party has won (based upon total number constituencies won)

```{r}
winning_rows = filter(parliament, Position==1)
summary_4 = winning_rows %>% group_by(YEAR, PARTY) %>% summarise(TOTAL=n()) %>% arrange(YEAR, -TOTAL) %>% group_by(YEAR) %>% summarise(first(PARTY))
summary_4 
```

### Question 5
Identify year wise percentage of winning for female candidates?

```{r}
summary_5 = parliament %>% filter(SEX=='F') %>% group_by(YEAR) %>% summarise(total=n(), won = length(which(Position==1))) %>% mutate(perc_win=won/total*100) %>% arrange(YEAR)
summary_5
```

